---
title: "General Social Survey"
author: "Carla Seals"
date: "May 11th 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

```



# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. First explain what you make out of comparing how people voted in 2012 with 2016.

How are the "voted in 2016" and "voted in 2012" related?

$H_0:$  There's no relations betweens the two years 2016 and 2012.

$H_A:$  Voting in the years 2016 and 2012 is related.

## Methods

The methods we used are categorical variables and chi-squared test to test the hypothesis that voting in 2016 is /is not independent of voting in 2012.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent", xlab="voted in 2016") 

```


This Bar Chart shows more people voted in 2012 than those that didn't vote in 2012.


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!--> 


```{r}
votetab=xtabs(~VOTE16+VOTE12, data=mydata4)
print(votetab)
```

The numerical summary show a cross tabulation between those that voted and did not vote in 2012 and 2016.


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(~VOTE16+VOTE12, data=mydata4)
fisher.test(votetab)
```


According the Chi-squared test and the Fisher's Exact Test the p-value is less than 0.05. The hypothesis clearly show the voters in 2016 are related to the voters in 2012.However, we reject the null hypothesis and accept the alternative. The Bar chart, The Numerical Summary were also used to show a clear relations between the people that voted in 2016 and 2012.

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)?

$H_0:$ Belief in the bigbang theory(BIGBANG) independent in the afterlife.

$H_0:$ Belief in the Big Bang is related to belief in the afterlife.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Using categorical variables and a Chi-squared test, We will test the hypothesis that belief in the Big Bang is or is not related to belief in the afterlife.



##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata4, type ="percent",xlab="Belief In The Big Bang")
```



#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
bangtab=xtabs(~BIGBANG+POSTLIFE, data=mydata3)
fisher.test(bangtab)
```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(~BIGBANG+POSTLIFE, data=mydata3)
          
```


```{r}
votetab=xtabs(~BIGBANG+POSTLIFE, data=mydata3)

fisher.test(votetab)
```
The hypothesis is that the belief of big bang and the belief in postlife is related. The p value is 0.0065, so we reject the null hypothesis since we dont have enough evidence to conclude that the two are related. If you believe in the big bang theory then you believe in post life.


# Introduction Question 3


<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This time compare the opinion on death penalty (CAPPUN) with opinion on courts (COURTS).

This data came from 2018 General social survey research question:

$H_0:$ opinion on death penalty (CAPPUN) IS INDEPENDENT OF THE OPINION ON COURTS (COURTS)  

$H_A:$ opinion on death penalty (CAPPUN) is related to the opinion on courts  

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
The methods we used are catigorical variables and chi-squared test to test the hypothesis that opinion on the death penalty theory is independent of the opinion on courts.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~CAPPUN+COURTS, data=mydata4, type =
"percent", xlab='view on CAPPUN', main='view on courts')
```


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
xtabs(~CAPPUN+COURTS, data=mydata4)
```

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(~CAPPUN+COURTS, data=mydata4)
```

```{r}
votetab=xtabs(~CAPPUN+COURTS, data=mydata4)
fisher.test(votetab)
```


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The hypothesis is that the opinion of the court is related to the opinion on death penalty and vice versa, the two variables are related based on the results of the chi-square test. 
We reject the null hypothesis since p is equal to 0. There is evidence to conclude that both the opinions are related. 